### 2016-10-17 result.tex
rough edits for youdin's comments

\section{
  Dusty vertical shear instability}\label{results}

%When $\tstop=0$ we may construct exact dusty equilibra by specifying
%the dust-to-gas ratio, as described in \S\ref{eqm}.

We present numerical solutions to the linearized
equations to examine the effect of dust on the VSI. 
We consider perfectly coupled dust ($\tstop=0$) so that the 
equilibria defined in \S\ref{eqm} are exact steady states.

For simplicity, we mostly consider constant midplane
dust-to-gas ratios $\epsilon_0$ and its characteristic height
$H_\epsilon$. The latter is specified via the dust layer thickness
$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
$H_\epsilon$ in \S\ref{varHd}.)   
These, together with the perturbation wavenumber $k$, comprises main the
parameters of the linear problem. We choose a power-law 
midplane gas density profile, $\rho_\mathrm{g0}\propto r^q$ with $p =
-1.5$. The power-law index for the 
temperature profile is $q=-1$; and we set the gas disk aspect-ratio
$h_\mathrm{g}=0.05$. These are fiducial values used by 
\citet[][\citetalias{lin15} in this section]{lin15}.

Two boundary conditions are needed for the $\tstop=0$
problem. This is expected since this problem is equivalent to
adiabatic hydrodynamics \citep[e.g.][]{lubow93}. We 
impose solid boundaries so that $\delta v_z(\pm\zmax)=0$.  
We solve the linearized equations as a generalized eigenvalue problem 
using a pseudo-spectral code adapted from \cite{lin15}. Amplitudes 
are expanded in Chebyshev polynomials up to order $N_z=512$. We check
results using Eq. \ref{vsi_check}.

\subsection{Qualitative expectations}\label{vsi_est}
\cite{lin15} found the appropriate way to compare 
destabilizing vertical shear to stabilizing vertical buoyancy is
$r\p_z\Omega/\OmK$ against $N_z^2/\OmK^2$. We see from \S\ref{vertshear}
and \S\ref{vbuoyancy} that $|r\p_z\Omega|\sim q h_\mathrm{g}\OmK$ for a
thin, gas dominated disk; while $N_z^2\sim \epsilon\OmK^2$. Thus we
expect dust-induced buoyancy forces to stabilize the disk against the
VSI where $\epsilon \gtrsim h_\mathrm{g}$.

%\begin{align}

\subsection{Effect of dust-loading}
We first vary the midplane dust-to-gas ratio 
$\epsilon_0\in[10^{-3},1]$, fixing the dust thickness to  
$\Hd=0.99\Hg$. Then  $\epsilon$ is roughly constant with height. We
set the vertical domain to $\zmax=5\Hg$.

Fig. \ref{compare_vshear_fixHd} compares the basic state
vertical shear rate, which is destabilizing, and the vertical buoyancy
frequency, which is stabilizing. For the nearly dust-free disk
$\epsilon_0=10^{-3}$ the vertical shear dominates buoyancy for all
$|z|>0$. However, a heavy dust-load with $\epsilon_0=1$ renders the 
buoyancy to dominate over vertical shear in the disk atmosphere 
$|z|\gtrsim 2.5\Hg$. We thus expect instability all heights for 
$\epsilon_0=10^{-3}$, but to be restricted to the midplane for
$\epsilon_0=1$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_fixHd} 
  \caption{Vertical shear rate (solid) compared to vertical buoyancy
    (dotted) in a locally isothermal, dusty disl with midplane dust-to-gas ratio
    of $\epsilon_0=10^{-3}$ (black) and $\epsilon_0=1$ (blue). 
    The dust layer thickness is fixed to $\Hd=0.99\Hg$ so the 
    dust-to-gas ratio $\epsilon$ is approximately constant with
    height. 
    \label{compare_vshear_fixHd}
    }
\end{figure}

Fig. \ref{vsi_dust_loading} show unstable modes for different values
of $\epsilon_0$ with fixed perturbation wavenumber  $k\Hg = 30$. The
eigenvalue distributions for $\epsilon_0 \leq 10^{-2}$ are similar to the
dust-free fiducial case considered by \citetalias{lin15}. This is
expected since $\epsilon_0 < \hgas$ (\S\ref{vsi_est}). Eigenvalues 
consists 
of the roughly horizontal `body modes' and the nearly-vertical
`surface modes'. The latter is due to the imposed vertical boundaries
\citep{barker15}.

We find that increasing the dust-to-gas ratio reduce VSI growth
rates. Notably, `surface modes', which are typically fastest growing in
the dust-free case, are surpressed in dusty disks for $\epsilon_0\geq
0.1$ (i.e. $\epsilon_0> \hgas$).  
The body modes' growth rates remain $ O(h_\mathrm{g}\OmK)$ 
but their oscillation frequency increases with
$\epsilon_0$. %i.e. increases with buoyancing freq 
The total number of modes do not change
significantly. This is in contrast with the effect of increasing
cooling times, where \citetalias{lin15} find fewer unstable modes.
%until the system is eventually completely stable.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_eigenvals_kx30Hd1} 
  \caption{Unstable modes in a locally isothermal, perfectly coupled
    dusty disk with fiducial parameters
    $(p,q,h_\mathrm{g}, \Hd/\Hg )=(-1.5,-1,0.05, 0.99)$. The real
    frequency $\omega$ and growth rates $s$ are shown for a range of
    midplane dust-to-gas ratios $\epsilon_0=\rho_\mathrm{g0}/\rho_\mathrm{d0}$. 
    \label{vsi_dust_loading}
    }
\end{figure}

The lowest frequency `fundamental' body mode is energetically dominant
because the entire disk column is perturbed \citep[cf. surface modes
  which only disturb the disk boundaries,][]{umurhan16c}. In Fig. \ref{vsi_dust_loading2d}
we compare the fundamental mode between the nearly 
dust-free case $\epsilon_0=10^{-3}$ an a dusty disk with
$\epsilon_0=1$. Dust-loading preferentially
stabilizes the disk atmosphere against the VSI, restricting
meridional motions to $|z|\lesssim 2\Hg$. 
This is consistent with Fig. \ref{compare_vshear_fixHd} comparing the basic state vertical
shear and buoyancy. Notice also the perturbed dust-to-gas ratio,
$\delta\epsilon$, shifts from being maximized at the disk boundaries
to the midplane.

\begin{figure}
%  \includegraphics[scale=0.54, clip=true, trim=0cm 2.5cm 0cm 0cm]{figures/result2d_dg1d-3.ps}\\
%  \includegraphics[scale=0.54]{figures/result2d_dg1.ps} 
\includegraphics[scale=0.32, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_dg1d-3.ps}\includegraphics[scale=0.32, clip=true, trim=1.8cm 0cm 0cm 0cm]{figures/result2d_dg1.ps}
  \caption{Fundamental dusty VSI mode in real space for midplane dust-to-gas
    ratio $\epsilon_0=10^{-3}$ (left) and $\epsilon_0=1$
    (right). The color scale shows the perturbation to the
    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
    \label{vsi_dust_loading2d}
    }
\end{figure}

In Fig. \ref{vsi_dust_loading_vareps} we plot the growth rates as a
function of $\epsilon_0$ for different perturbation wavenumbers $k$. We
find that dust-loading stabilizes the VSI more effectively for shorter
wavelength perturbations. This is because for high wavenumbers the
dominant modes are surface modes, which are effectively stabilized by
dust-loading since bouyancy forces are larges near the disk
boundaries. The figure suggest that VSI becomes much less efficient
for $\epsilon_0\gtrsim 0.1$ and $k\Hg\gtrsim 50$.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_eigenvals_vareps2} 
  \caption{Maximum growth rate of the dusty VSI as a function of the
    midplane dust-to-gas ratio $\epsilon_0$ for perturbations with
    different radial wavenumbers $k$. The dust layer thickness is
    fixed to $\Hd\simeq \Hg$. 
    \label{vsi_dust_loading_vareps}
    }
\end{figure}

\subsection{Effect of dust thickness} 
We now vary the dust layer thickness $\Hd$ but fix the metalicity 
$Z \equiv \epsilon_0 \Hd/\Hg = 0.03$. This then gives the
required midplane dust-to-gas ratio. Since we will consider thin dust
layers, here we use a smaller 
domain with $\zmax=2\Hg$ so that $\epsilon$ does not become
too small.

We analyze two disks with $\Hd=0.1\Hg$ and 
$\Hd=0.99\Hg$. Fig. \ref{compare_vshear_fixZ} compares the vertical
vhear rate and buoyancy frequency. For $|z|\gtrsim 0.4\Hg$ the two
disks have the same profile with vertical shear dominating over
buoyancy. We thus expect perturbations away from the disk midplane in 
both cases. For $|z|\lesssim 0.4\Hg$, a thin dust 
layer with $\Hd=0.1\Hg$ boosts the vertical shear rate, but the
associated buoyancy is larger still, implying the mid-plane should be
stable. 
%Thus we expect the mid-plane of 
%the disk with $\Hd=0.1\Hg$ to have limited vertical motions.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_fixZ} 
  \caption{Vertical shear rate (solid) compared to vertical buoyancy
    (dotted) in a locally isothermal, dusty disk 
    with metalicity $Z=0.03$ and dust thickness $\Hd=0.1\Hg$
    (black) and $\Hd=0.99\Hg$ (blue). 
    \label{compare_vshear_fixZ}
    }
\end{figure}

Fig. \ref{result2d_fixZ} compares the fastest growing VSI body modes  
with $k\Hg=30$ for the two cases above.\citepalias[The thinner domain
  adopted here eliminates surface modes, ][]{lin15}.   
We find very similar mode 
frequencies 
\begin{align*}
  \sigma = \begin{cases}
    \left(0.3053\ii - 0.8142\right)h_\mathrm{g}\OmK & \Hd=0.99\Hg, \\
    \left(0.3178\ii - 1.2237\right)h_\mathrm{g}\OmK & \Hd=0.1\Hg,
  \end{cases}
\end{align*}
since the vertical shear profile is similar throughout most of the
disk. However, meridional motions are surpressed near the midplane of
the $\Hd=0.1\Hg$ disk, as expected from the larger buoyancy frequency
as compared to vertical shear there. This leads to a structure
analogous to PPD dead zones: a quiescent midplane between active
surface layers \citep{gammie96}.

\begin{figure}
  \includegraphics[scale=0.32, clip=true, trim=0.5cm 0cm 3cm 0cm]{figures/result2d_Hd1.ps}\includegraphics[scale=0.32, clip=true, trim=1.8cm 0cm 0cm 0cm]{figures/result2d_Hd0d1.ps} 
  \caption{Fastest-growing dusty VSI mode in real space for midplane
    dust layer thickness $\Hd=0.99\Hg$ (left) and $\Hd=0.1\Hg$
    (right). The dust content is fixed to
    $\Sigma_\mathrm{d}=0.03\Sigma_\mathrm{g}$. 
    The color scale shows the perturbation to the
    dust-to-gas ratio, $\delta\epsilon$; and the arrows show
    $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$.
    \label{result2d_fixZ}
    }
\end{figure}

Fig. \ref{compare_eigenvals_fixZ} shows the maximum VSI growth rates
as a function of $\Hd$. As before, we find growth
rates are most affected by the vertical structure of the dust layer
when the perturbation wavenumer is large. Notice VSI growth rates converge as
$\Hd\to 0$. %to dust free values?    
Thus a thin dust layer, however large its associated vertical shear,
does not affect the VSI. The non-monotonic behavior for $\Hd\gtrsim
0.5\Hg$ arises because the vertical buoyancy frequency 
\begin{align*}
N_z^2(H_d;z,Z) \simeq &Z\Hg z^2
\exp{\left(-\frac{z^2}{2\Hg^2}\right)}\OmK^2\notag\\
&\times 
\frac{1}{\Hd}\left(\frac{1}{\Hd^2} -
\frac{1}{\Hg^2}\right)\exp{\left(-\frac{z^2}{2\Hd^2}\right)}  
\end{align*}
is a non-monotonic function of $\Hd$ at fixed $z$. At $z=\Hg$ and $z=2\Hg$
the buoyancy frequency is maximized for $\Hd\simeq0.5\Hg$ and
$\Hd\simeq 0.8\Hg$, respectively. This is consistent with the abscissa
of minima in growth rates in Fig. \ref{compare_eigenvals_fixZ}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_eigenvals_fixZ} 
  \caption{Maximum VSI growth rate for different perturbation
    wavenumbers $k$ as a function of the dust layer
    thickness $\Hd$ at fixed metalicity $Z=0.03$. 
    \label{compare_eigenvals_fixZ}
    }
\end{figure}

\subsection{Axisymmetric stability of ultra-thin dust layers}
In \S\ref{iso_perfect} we asserted that in the limit of perfectly
coupled dust (as considered here) that radially smooth, strictly
isothermal disks are stable against axisymmetric perturbations, no
matter how thin the dust layer is. We now demonstrate this
numerically.

For comparison with previous studies of dusty disks, we also use 
the Richardson number $\rich \equiv N_z^2/\left(r\p_z\Omega\right)^2$
to label the calculations here \citep{youdin02}.   
Numerical simulations show   
non-axisymmetric instabilities develops when $\rich\lesssim0.1$, brought about
by very thin dust layers \citep{chiang08, lee10}. We show that \emph{axisymmetric}
instabilities never develop, however small $\rich$.

We shall consider ultra thin dust layers with $\Hd\leq 0.01\Hg$ and thus
restrict the vertical domain to $\zmax = 0.02\Hg$. We fix the 
metalicity $Z=0.01$ so the midpane dust-to-gas ratio $\epsilon_0$ is 
$O(1)$.  We consider radial wavenumbers such $k\Hd=1$.

Fig. \ref{ultra_thin} shows the maximum growth rate  as a function of
the radial temperature gradient, $q$. For all cases the vertical
shear is dominated by that due to the dust layer (see
\S\ref{vertshear}). However, we see that growth rates $s\propto
|q|$. In particular, this holds for $\rich < 0.1$, the critical value
for non-axisymmetric instabilities.

Axisymmetric instability is associated with the thermal
contribution to vertical shear: as $q\to0$,  $\p_z\Omega$ becomes
entirely due to dust-settling and there is no instability ($s\to
0$). This result is independent of $\rich$, so the Richardson number
is not useful in characterizing the axisymmetric stability of dust
layers.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_eigenvals_thindust} 
  \caption{Maximum VSI growth rate for ultra-thin dust layers $\Hd
    \leq 0.01\Hg$. Vertical shear is dominated by that due to
    dust-settling, but axisymmetric instability is associated with the
    radial temperature gradient $q$.
    \label{ultra_thin}
    }
\end{figure}

\subsection{Effect of a radially-varying dust-to-gas ratio}\label{varHd}
{\bf move to appendix?}

Here we allow the dust-to-gas ratio $\epsilon$ to depend on
$r$. Specifically we let $\epsilon_0\propto r^{-1}$ and
$H_\epsilon \propto \Hg$ (cf. constant values in the previous
calculations). Then
\begin{align*}
  \frac{\p\epsilon}{\p r} =
  \left(\frac{z^2}{H_\epsilon^2}\frac{d\ln{\Hg}}{dr} -
    \frac{1}{r}\right)\epsilon.   
\end{align*}
We fix $Z=0.01$, $\Hd=0.8\Hg$ at $r=r_0$, and
$k\Hg=30$.

Fig. \ref{compare_vshear_varHd} shows that a radially-varying
dust-to-gas ratio increases the magnitude of the vertical shear rate
away from the midplane (see also Eq. \ref{vshear2}). Thus we find
higher VSI growth rates, as shown in Fig. \ref{vsi_dust_varHd}. 
Surfaces modes are more effectively enhanced by the additional vertical 
shear induced by $\p_r\epsilon$. However, all growth rates remain
$O(\hgas\OmK)$. Importantly, the increase in the growth rate of the
fundamental mode is small. Thus we do not expect the radial
dependence in $\epsilon$ to significantly affect the VSI in a dusty
disk.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_vshear_Nz2_varHd} 
  \caption{Vertical shear rate (solid) compared to vertical buoyancy
    (dotted) in a locally isothermal, dusty disk 
    with metalicity $Z=0.01$ and dust thickness $\Hd=0.8\Hg$.
    Black curves assume a dust-to-gas ratio that only depends on
    height; whereas the blue curve also allows a radial dependence in
    $\epsilon$. See text for details.  
    \label{compare_vshear_varHd}
    }
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_eigenvals_varHd} 
  \caption{Unstable VSI modes in the disk models of
    Fig. \protect\ref{compare_vshear_varHd}.  
    \label{vsi_dust_varHd}
    }
\end{figure}

\subsection{Validity}
We considered $\tstop=0$ in order to perform stability analysis on an
exact steady state. When $\tstop\neq0$, a 
disk initialized as in \S\ref{eqm} will evolve 
due to dust-gas relative drift. In particular, in the
absence of turbulent/diffusive processes, dust particles settle to the
midplane on a timescale $t_\mathrm{settle}\sim 1/\tstop\OmK^2$
\citep{takeuchi02}. However, we expect the above results to remain
valid provided the VSI growth
timescale $t_\mathrm{grow}\ll t_\mathrm{settle}$. Since 
VSI growth rates are $O(h_\mathrm{g}\OmK)$, we require 
\begin{align}
  \tstop\OmK \ll h_\mathrm{g}. 
\end{align} 
For thin PPDs with $h_\mathrm{g}\sim 0.05$, this is satisfied 
for $\tstop\OmK \ll O(10^{-2})$.