### 2017-05-09 paper.tex
misc edits to the main text overall

%\documentclass[12pt]{aastex}
\documentclass[iop, numberedappendix]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
%\usepackage{paralist}
\shorttitle{Isothermal dusty gas dynamics}
\shortauthors{M-K.\ Lin, A.N. Youdin}

\usepackage{amsmath}
\usepackage{paralist}
\usepackage{url}
\usepackage{bm}
%\usepackage{hyperref}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\zmax}{z_\mathrm{max}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\khat}{\hat{k}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\rich}{\mathrm{Ri}}
\newcommand{\Htilde}{H_\epsilon}
\newcommand{\Hdust}{H_\mathrm{d}}
\newcommand{\Hgas}{H_\mathrm{g}}
\newcommand{\hgas}{h_\mathrm{g}}
\newcommand{\tepsilon}{f_\mathrm{d}}
\newcommand{\rhod}{\rho_\mathrm{d}}
\newcommand{\rhog}{\rho_\mathrm{g}}
\newcommand{\tstop}{t_\mathrm{s}}
\newcommand{\taus}{\tau_\mathrm{s}}
\newcommand{\w}{ \widetilde{W}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\Hg}{H_\mathrm{g}}
\newcommand{\Hd}{H_\mathrm{d}}
\newcommand{\seff}{S_\mathrm{eff}}
\newcommand{\tcool}{t_\mathrm{cool}} 
\newcommand{\actaa}{Acta Astron.}
\defcitealias{lin12}{L12} 
\defcitealias{lin15}{LY15}
\defcitealias{nelson13}{N13}
\defcitealias{barker15}{BL15}
\defcitealias{mcnally14}{MP14}
\defcitealias{goldreich67}{GS67}

\def \OmK {\Omega_{\rm K}}

\begin{document}

%\title{On the entropy of 
%  isothermal dusty gas: application to protoplanetary disks}  
\title{A thermodynamic view of dusty protoplanetary disks}
 % \title{Dust-free modeling of dusty protoplanetary disks I:
%axisymmetric stability}
\author{Min-Kai Lin\altaffilmark{1,2,3}}
\author{Andrew N. Youdin\altaffilmark{1}}
%\affil{Department of Astronomy and Steward Observatory, 
%University of
 % Arizona, Tucson, AZ 85721, USA\\
%Institute of Astronomy and Astrophysics, Academia Sinica, 
%Taipei 10617, Taiwan}
\altaffiltext{1}{Department of Astronomy and Steward Observatory, 
University of
Arizona, Tucson, AZ 85721, USA}
\altaffiltext{2}{Steward Theory Fellow}
\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
Taipei 10617, Taiwan
}
\email{mklin@asiaa.sinica.edu.tw}

\begin{abstract}
  Small solids embedded in gaseous protoplanetary disks are subject
  to strong dust-gas friction. Consequently, tightly-coupled dust
  particles almost follows the gas flow. 
  This is analogous to entropy conservation
  following an ideal gas.   
  Motivated by this similarity, we develop a thermodynamic 
  framework to study dusty gas dynamics in protoplanetary disks. We  
  show that an isothermal dusty gas behaves like an  
  adiabatic pure gas; and that finite dust-gas coupling may be
  regarded as an effective 
  heating/cooling. We exploit this correspondence 
  to deduce that   
\begin{inparaenum}[1)] 
\item 
 perfectly coupled, thin dust layers cannot cause axisymmetric 
 instabilties; 
\item 
  the streaming instability neccessarily involves a gas pressure
  response 
  that lags behind dust density; 
%  the streaming instability neccessarily involves a phase lag between
%  gas pressure and dust densities; 
\item dust-loading introduce buoyancy forces that generally 
  stabilizes the vertical shear 
  instability associated with global radial temperature gradients. 
\end{inparaenum}  
%growth rates, and 
%preferentially stabilize modes with small radial 
%wavelengths, as well as `surface modes' concentrated near the
%disk boundaries. 
%These results can be understood in terms of 
%buoyancy forces in an isothermal dusty disk. 
%We show that vertical shear arising from
%dust settling does not lead to axisymmetric instability. 
We also discuss future applications of this thermodynamic framework in
finding dusty analogs of other hydrodynamic 
phenomena (e.g. Rossby wave instability, convective overstability, and 
zombie vortices), in simulating dusty protoplanetary disks
using purely hydrodynamic codes, as well as potential generalizations
of the dusty/adiabatic gas equivalence principle.

%We briefly describe extending this
%equivalence principle to polytropic and adiabatic gas.

\end{abstract}

\section{Introduction}
%si, sgi as examples
Protoplanetary disks are comprised of a mixture of gas and dust
\citep{chiang10}. Although the overall dust content is small
($\sim 1\%$ by mass), their presence can have profound impacts on the 
gas dynamics and vice versa. Dust-gas friction introduces phenomena
that are absent in pure gas disks. Important examples include the
streaming instability 
\citep[SI,][]{youdin05a,youdin07b,johansen07}, secular gravitational
instabilities \citep[SGI,][]{ward00,youdin11,michikoshi12,takahashi14} and
Kelvin-Helmholtz instabilities
\citep{chiang08,barranco09,lee10}. These lead to structure and
planetesimal formation, as well as turbulence. 
More recently, new 
dusty instabilities have also been uncovered through numerical
simulations \citep{loren15,loren16,lamb16}, but these are not yet well
understood.

%particles 
Current state-of-the-art models of dusty protoplanetary
disks directly simulate gas dynamics coupled to explicit
Lagragian dust particles
\citep{nelson10,bai10,johansen11,yang14,zhu14,gibbons15,simon16,baruteau16}. 
The equation of motion for each particle is solved 
including dust-gas drag; the strength of which is measured 
by a stopping time $\tstop$ --- the decay timescale for the 
relative velocity between gas and dust. Thus small $\tstop$
corresponds to tightly-coupled particles.

%refs

Another approach is to model the dust population as a continous, pressureless
fluid 
\citep{paardekooper06b,meheut12,laibe12,loren14,fu14b,surville16}. The 
hydrodynamic equations are evolved for two fluids: the gas and dust with density and
velocity $\rhog$, $\bm{v}_\mathrm{g}$ and $\rhod$, 
$\bm{v}_\mathrm{d}$, respectively. Source terms are introduced into
the momentum/energy equations to model dust-gas drag. This approach
can take advantage of existing numerical methods/codes for
simulating pure gas dynamics.

%refs

It is possible to re-formulate the two-fluid equations into that for a
single fluid with total density $\rho$ and center-of-mass
velocity $\bm{v}$ of the dusty gas mixture
\citep{laibe14}. The standard hydrodynamic equations for
$\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
the dust-to-gas ratio $\rhod/\rhog$ and the relative dust-gas velocity
$\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$. The latter also contributes to 
additional source terms in the momentum and energy equations.

The one-fluid approach is appropriate for 
simulating gas tightly coupled to dust particles with small $\tstop$.  
This can be numerically challenging for two-fluid models and 
following Lagragian particles; but is straight forward with the 
one-fluid formulation \citep{price15}.  The single fluid framework 
also has the important advantage that it is 
amenable to standard analytical methods \citep{youdin05a,jacquet11}.  
This is the necessary approach to gain physical insight.

In this work, we are motivated to develop a general analytical
framework to study the stability of dusty protoplanetary 
disks by drawing on the physical and mathematical parallel between
dusty gas and standard hydrodynamics. This is possible in certain limits of astrophysical
interest, allowing us to interpret dusty phenomena through standard
hydrodynamics.

We consider tightly coupled dust particles with $\tstop\Omega\ll
1$, where $\Omega$ is the disk orbital frequency. We may then use the
`terminal velocity approximation' to set the dust-gas relative drift
to $\Delta \bm{v} = \tstop\nabla
P/\rhog$, where $P$ is the gas pressure \citep{youdin05a, 
  jacquet11,laibe14}. This approximation reflects the physical result
that dust particles tend to drift towards pressure maximum.

 We also consider irradiated protoplanetary disks 
 in which the temperature is set externally
\citep{chiang97,stam08}. This allows us to adopt a locally isothermal
equation of state, $P = c_s^2\rhog$, where $c_s$ is a prescribed
sound-speed profile. This physically represents short cooling times
\citep{lin15}.

Under these two approximations, we show that the equations of 
isothermal dusty gas dynamics are equivalent to that of adiabatic, pure gas
dynamics. This is because tightly-coupled dust particles are almost
entrained in the gas, which is analogous to the advection
of entropy following an adiabatic fluid. We show that the   
effective entropy of isothermal dusty gas scales with its gas
fraction. Finite dust-gas drag act as effective heating/cooling: the relative
drift between dust and gas causes a fluid parcel to exchange dust
particles with its surroundings; an effect similar to heat exchange
between a gas parcel and its surroundings.

%suitable for 
%analytic studies 
%more importantly, can understood

By re-casting the dusty problem into standard  
hydrodynamics, we can apply established methods to generalize previous stability 
results for pure gas dynamics to dusty disks.  
We show that strictly isothermal protoplanetary disks with
perfectly-coupled dust are generally 
stable to axisymmetric perturbations.%, regardless of how  thin the dust
%layer is. 
{\bf should we mention vertically well mixed case here?} 
We put forward a thermodynamical interpretation of 
dust-gas drag instabilities, and apply it to the streaming instability.   
We also study the effect of dust-loading on the vertical shear
instability previously considered in pure gas disks
\citep[VSI,][]{nelson13,lin15,barker15}.

This paper is organized as follows. We first present, without
derivation, the correspondence between dusty gas and     
ideal hydrodynamics in \S\ref{key_idea}. In \S\ref{grow_osc} 
we discuss a key idea in fluid dynamics, relevant to our 
discussion on  SI and VSI: instabilities associated with
pressure-density phase lags.   
%This means that dusty-gas phenomena, such 
%as SI and VSI, have thermodynamical interpretations.

We begin our formal discussion in \S\ref{setup} by listing the 
basic equations describing tightly-coupled dusty gas. We show that for a
fixed equation of state, these are in 
fact the usual hydrodynamic equations in disguise. 
We describe equilibrium solutions in \S\ref{eqm} that are used in
explicit calculations. In \S\ref{limits} we discuss 
general stability properties of dusty disks. We then move onto
explicit stability analyses. In  
\S\ref{linear_problem} we list the linearized equations and show that 
they reproduce previous results for SGI and SI. In \S\ref{results} we
extend the VSI to dusty disks. We   
discuss future applications of the dusty/adiabatic gas equivalence in 
\S\ref{discussion} before summarizing in \S\ref{summary}.

\section{Dust-free modeling of dusty protoplanetary disks}\label{key_idea}
{\bf is this section needed?} 
The key concept we introduce in this work is the description
of dusty protoplanetary disks \emph{without} explicit dust
equations. This is best illustrated by writing down our model   
equations for a mixture of isothermal gas and small
dust particles: 
\begin{align}  
  &\frac{\p \rho}{\p t} + \nabla\cdot\left(\rho\bm{v}\right)= 0, \label{mass0}\\ 
  &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} = -
  \frac{1}{\rho}\nabla  P - \nabla \Phi_*,\label{mom0}\\ 
 & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P 
  = - P \nabla\cdot\bm{v}
                + \mathcal{C}(\rho, P; \tstop)\label{en0}, 
\end{align}
where $\rho=\rhod + \rhog$ is the total density, $P$ is the pressure,
$\Phi_*$ is the stellar gravitational potential,   
and $\mathcal{C}$ represent effects of dust-gas drag, which
is paremterized by the stopping time $\tstop$ ($\mathcal{C}=0$ for
$\tstop=0$). Similar equations for dusty-gas are given by
\cite{price15}.

Eq. \ref{mass0}---\ref{en0} appear identical to adiabatic gas
dynamics, even though they physically represents a two-phase,
dusty-gas. The main difference is the term $\mathcal{C}$. Here it does
not represent physical cooling, as the system is isothermal by
assumption. Instead, dust-gas drag causes particles to drift, hence by
momentum conservation a gas drift in the opposite direction, which
changes the pressure distribution.

Since the above equations are similar to ideal gas dynamics, it 
means that many of the results derived for ideal hydrodynamics can be  
generalized to dusty disks without additional effort. This includes 
the important consequence of phase lags between density and pressure,
discussed next.

\section{Growing oscillations by doing work}\label{grow_osc}
A general result in fluid dynamics is that work is done whenever
oscillations in the pressure and density of the fluid are not in
phase. If  the average work done is positive, then oscillation
amplitudes grow: the work done allows the system to `overshoot' beyond 
the amplitude of preceeding oscillations.

%To see this, we first compute the work done
%assuming periodic oscillations. We then show that if the work done is positive, then the
%oscillation amplitude in fact grows: the work done leads the
%system to `overshoot' past the amplitude of the preceeding oscillation
%cycle. 
%its equilibrium. {\bf should be overshoot the max displacement}

Fig. \ref{pdv_cartoon} gives a graphical demonstration that 
if pressure lags behind density, then positive is work done 
because it leads to a clockwise path in the 
`P-V' plane. The annotations consider the particular case where the
phase lag arises because the fluid has two components (gas and dust,
see below) but the following description is general.

From $A$ to $B$, a fluid parcel is expanding to return to 
equilbrium, but pressure is still 
increasing. This over-compensation will cause the parcel to expand
beyond the maximum volume of the previous cycle. 
Similarly, from $C$ to $D$ the parcel is already contracting towards
equilibrium, but pressure 
is still dropping. This allows the  
parcel's contraction to over-shoot the maximum density attained in the
previous cycle. 
The overall positive work done leads to growth in the oscillation amplitude. 
This effect is similar to being pushed downwards on a swing when 
descending.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/drag}\\
\includegraphics[width=\linewidth]{figures/pdv}
  \caption{Thermodynamic interpretation of growing oscillations in a
    fluid. Top: pressure and density evolution in time. 
    Bottom: oscillation cycle in the P-V plane. In the case shown, 
    dust-drag causes 
    pressure to lag behind density, which results in a clockwise path in
    the P-V plane, implying positive work done by the fluid, which 
     would increase oscillation amplitudes.  
    \label{pdv_cartoon}
  }
\end{figure}

There are several situations where the pressure and density of a fluid
are not in phase. The obvious case is if the fluid is subject to
external heating/cooling. For example, in strongly irradiated
protoplanetary disks the
disk temperature  $T(r)$ is time-independent but varies with the cylindrical radius
$r$ from the star \citep{chiang97}. Since $P\propto \rhog T$ for an
ideal gas, %the pressure of a gas parcel varies with  
%position as well. 
there is no reason to expect $P$ and $\rhog$ to 
be in phase as a gas parcel oscillates between different radii. In
fact, we will show that this is a fundamental property of the vertical
shear instability \citep{lin15}.

%temperature of a fluid parcel changes with position

Another possibility, as annotated in Fig. \ref{pdv_cartoon}, is a dusty gas. The
relevant density here is the total density $\rho = \rhod + \rhog$, 
but the fluid pressure $P$ is due to gas only. If dust were 
perfectly-coupled to the gas, then $P$ and $\rho$ would be in phase, 
and there is no work done. However, for finite dust-gas drag, $\rhod$ and
$\rhog$ are not neccessarily in phase, because dust particles can
drift relative to the gas.   
\emph{If} this causes $P$ to lag 
$\rho$, then the positive work done would lead to growing
oscillations. Indeed, we show that this is true for the streaming
instability.

In order to apply this thermodynamic interpretation of dust-gas drag
instabilities, we need to develop a formal analogy between dusty-gas
and pure gas dynamics. We show this is possible in the limit of
strong drag and a fixed gas equation of state.

\input{model}
\input{linear}
\input{result}
\input{summary}

%{\bf acknowledgements. s. inutsuka for asking about kz=0 limit}

\appendix
\input{appendix}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}