### 2017-05-09 summary.tex

%growth rates not sig changed 
%v shear from thin dust layers are overwhelmed by 
\section{Discussion}\label{discussion}

\subsection{Generalization to locally polytropic disks}\label{gen_poly}

%\subsubsection{Locally polytropic disks}
We can extend the dusty/adiabatic gas correspondence to 
other fixed equations of state. As an example, consider the locally
polytropic disk 
\begin{align}
  P = K(r,z)\rhog^{\Gamma}, 
\end{align}
where $K$ is a prescribed function and $\Gamma$ is the constant
polytropic index. Then eliminating $\tepsilon$ from the dust equation
\ref{dusteq} gives 
\begin{align}\label{poly_energy}
  \frac{D P}{D t} &= - \Gamma P\nabla\cdot\bm{v}  + P \bm{v}\cdot\nabla
  \ln{K} + \mathcal{C},\\
  \mathcal{C}& = \frac{\Gamma P}{\rhog}\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right).
\end{align}
Thus the dusty gas behaves like a pure gas with adiabatic index
$\Gamma$. The entropy is given by 
\begin{align}
  \seff = \ln{\left[K^{1/\Gamma}(r,z)\left(1 - \tepsilon\right)\right]}.  
\end{align}
The results of \S\ref{iso_perfect}---\ref{dust_work} remain valid for the strictly 
polytropic disk with $K=$constant, while one sets $c_s^2\to K$ in
\S\ref{dusty_vsi_int}. 
%When $K$ is constant, the corresponding Solberg-Hoiland criteria for
%axisymmetric stability may be obtained from that in standard adiabatic 
%hydrodynamics \citep[e.g.][]{tassoul78} by
%using this definition of entropy in
%Eqs. \ref{dusty_solberg1}---\ref{dusty_solberg2}.

%In particular, we will explore whether this can surpress the suprious
%modes discussed in Appendix \S\ref{spurious_epi}.

%stratified case require more boundary conditions

\subsection{Dusty analogs of other gaseous instabilities}

%gi
%rwi

\subsubsection{Classic gravitational instabilities} %vertical structure 
The addition of dust enhances gravitational
instability (GI) because dust particles contribute to the
total disk mass but not thermal pressure, which effectively lowers the
disk temperature \citep[][]{thompson88,shi13}. For typical dust-loading 
$\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
large (e.g. due to dust settling) then the reduced temperature
$\widetilde{T}$ may be lowered to enable instability.

As noted in \S\ref{loc_iso_eos}, dust settling causes 
$\widetilde{T}$ to \emph{increase} away from the midplane. This contrasts
to previous studies of GI in vertically stratified disks
 where the temperature decreases
from the midplane \citep[e.g.][]{mamat10, kim12,lin14c}. 
While we expect only the total surface density and
characteristic temperatures are relevant to stability  
\citep{toomre64}, a non-trivial vertical temperature
structure, induced by dust, may modify the vertical structure of 3D
waves and unstable modes.

Another potential connection to previous results for gas disks is  
gravito-turbulence. Cooling, self-gravitating gaseous disks 
sustain a turbulent state where shock heating due to gravitational 
instabilities are balanced by radiative cooling 
\citep{gammie01}. Since dust-gas drift appears as a diffusion or 
cooling in our framework (Eq. \ref{eff_energy}), it may be conceivable
to have `dusty gravito-turbulence'. As self-gravity increases the
local density, the associated pressure maxima attract dust-particles,
but the back-reaction onto the gas may try to flatten the pressure
bump \citep{taki16}.

%see also
%\S\ref{loc_iso_eos}
%We thus expect the standard Toomre
%parameter to be modified to $Q = c_s\sqrt{1-\tepsilon}\Omega/\pi
%G\Sigma$.

\subsubsection{Rossby wave instability}
The Rossby wave instability (RWI) is a non-axisymmetric, 2D shear
instability that operates in thin disks when it has radial structure
\citep{lovelace99,li00}. These studies consider adiabatic pure gas and 
show instability is possible if there is an extremum in the generalized
potential vorticity 
%\begin{align}
 $\mathcal{V}_\mathrm{g} =
 \kappa^2\mathcal{S}_\mathrm{g}^{-2/\gamma}/2\Omega\Sigma_\mathrm{g}$, 
%\end{align} 
where $\mathcal{S}_\mathrm{g} = P/\Sigma_\mathrm{g}^\gamma$. Here $P$
should be interpreted as the vertically integrated pressure. The
non-linear result of the RWI is vortex formation \citep{li01}.

The dusty/adiabatic gas analogy imply the condition for RWI in a
polytropic dusty gas disk is an extremum   
\begin{align}
  \mathcal{V} =
  \frac{\kappa^2}{2\Omega\Sigma}\frac{1}{K^{2/\Gamma}}\left(1+
    \frac{\Sigma_\mathrm{d}}{\Sigma_\mathrm{g}}\right)^2.    
\end{align}
Thus RWI may also be triggered by extrema 
in the dust-to-gas ratio, e.g. narrow dust rings/gaps. This may lead 
to direct formation of dusty vortices, as opposed to dust-trapping by  
a pre-existing gas vortex \citep{barge95,lyra13}.

%{\bf so may it's not vortex formation then dust-trapping. could just
%  form dusty vortices in one go}

%gap edges. thin dust rings.

\subsubsection{Convective overstability}
The `convective overstability' (ConO) was discovered in pure gas, non-adiabatic, unstratified disk models    
where the radial buoyancy frequency is such that $N_r^2\equiv
F_r\p_rS/C_P  < 0 $ and cooling times $t_\mathrm{cool}\sim \OmK^{-1}$. This
combination leads to growing epicycles
\citep{klahr14,lyra14,latter16}.     
%with 
%growth rate $\propto  -N_r^2$ 
% Indeed, the effective radial buoyancy frequency in our dusty disk 
% with negligble background density gradient is unstable, 
% \begin{align}
%   N_r^2 \equiv -\frac{1}{\rho}\frac{\p P}{\p r}\frac{\p S_\mathrm{eff}}{\p r} =
%   -\frac{\rho}{P}F_r^2<0.  
% \end{align}

%The disk structure required for ConO might be realized in a locally
%isothermal dusty disk.

Now consider a strictly isothermal dusty disk where the relevant 
entropy is dust-induced, $\seff =-\ln{(1+\rhod/\rhog)}$. 
%The dusty/adiabatic gas
%analogy implies the replacement $S\to\seff = -\ln{(1+\rhod/\rhog)}$ in
%$N_r$. 
Since $F_r\propto -\p_rP > 0$ in typical disk models, ConO would
require $\p_r\seff<0$, implying the dust-to-gas ratio increases
\emph{outwards}. This might be realized at special radial locations in 
protoplanetary disks, such as planet gaps. Dust-settling itself may
also lead to the mid-plane dust-to-gas ratio to increase outwards
\citep{takeuchi02}.

%fact, the typical disk models used to study the streaming
%instability have zero background radial density gradient, in which
%case  $N_r^2= -\rho F_r^2/P < 0$.

Isothermal disks have $\tcool=0$, so the pure gas ConO cannot
exist. However, we have shown that dust-gas drag provides an effective
energy source/sink to the mixture (\S\ref{energy_analogy}). It would be
interesting to explore whether or not dust-gas drag can play the role
of cooling to enable a `dusty convective 
overstability'.

%Furthermore, 
%the effective radial buoyancy frequency in our dusty disk  
%with negligble background density gradient is unstable, 
%\begin{align}
%   N_r^2 \equiv -\frac{1}{\rho}\frac{\p P}{\p r}\frac{\p S_\mathrm{eff}}{\p r} =
%   -\frac{\rho}{P}F_r^2<0.  
% \end{align}

% One might then think of the streaming instability as a dusty relative
% to the convective  overstability. Explicit solutions in the small
% $\tstop$ limit show that streaming instability growth rates are
% also $\propto |N_r^2|$  \citep{jacquet11}, as is for the convective
% overstability \citep{lyra14,latter16}. Furthermore, both instabilities require
% finite effective cooling times, and both require $k_z\neq0$. 
% %{\bf note: si osc freq not too small compared to kappa}

% However, the convective overstability is most unstable at $k_x=0$
% \citep{lyra14}  whereas the streaming instability growth rate is maximized at finite $k_x$
% \citep{youdin05a}.

\subsubsection{Zombie vortices}

The `zombie vortex instability' (ZVI) is a non-axisymmetric,
non-linear instability discovered in pure gas disks
\citep{marcus15,umurhan16d}. It operates by having finite-amplitude perturbations
exciting `critical layers',  where the intrinsic wave frequency of the
perturbation matches the local vertical buoyancy frequency 
\citep{marcus13}. These critical layers roll 
up into vortices, exciting further critical layers. The process
repeats itself and the result is an array of vortices.

Since the necessary physical ingredient for the ZVI is vertical
buoyancy, it requires an almost purely adiabatic gas ($\tcool\OmK\gg
1$). This limits its applicability in typical protoplanetary disks to 
$\lesssim 1$AU \citep{lesur16, malygin17}. However, this consideration
assumes a dust-free disk as far as the dynamics is concerned.

On the other hand, we have show that dust-loading induces an effective
buoyancy even in isothermal disks. It is thus natural to ask whether or not
the required adiabatic conditions for ZVI can be realized through
dust-loading, and thus produce `dusty ZVI'. This may allow the zombie
vortices to develop $\gtrsim 1$AU in PPDs.

\subsection{Implications for numerical simulations}
The dusty/adiabatic gas equivalence means that pure
gas dynamics codes can be applied to simulate locally
isothermal/polytropic dusty gas. One need not implement a separate 
equation for the dust component: 
%In fact, no
%modification is needed  
%for strictly isothermal or polytropic gas perfectly-coupled to dust:  
 the standard energy equation substitutes for the dust continuity
equation. When $c_s$ (or $K$) is not constant and/or $\tstop\neq0$ one
should add corresponding source terms in the energy equation
(e.g. Eq. \ref{poly_energy}). 
The source term associated with dust-gas drag, $\mathcal{C}$, is 
analogous to radiative diffusion \citep{price15} or thermal
conduction, which is also common in modern hydrodynamics codes.

We have taken advantage of the dusty/adiabatic gas equivalence to 
convert the popular
\textsc{Pluto}\footnote{\url{http://plutocode.ph.unito.it/}}      
hydrodynamics code \citep{mignone07} into a dusty gas dynamics code appropriate for
simulating protoplanetary disks coupled to small dust. We will apply 
our modified code --- d\textsc{Pluto} --- to the streaming instability
and disk-planet interactions in an upcoming publication (Lin et al.,
in preparation).

\subsection{Higher order corrections}
{\bf notes for self. maybe exclude in final. maybe move into appendix}

The simplified one-fluid formulation, 
Eqs. \ref{masseq}---\ref{tempeq}, can 
produce spuriously-growing modes with frequency $|\sigma|\gtrsim
\Omega$ (see Appendix \S\ref{spurious_epi}). This 
results from neglecting terms of $O(|\widetilde{\bm{v}}|^2)$ after converting
the full two-fluid equations to the one-fluid formulation. The full
one-fluid equations 
%is significantly more complex than our 
%model \cite[e.g.][]{laibe14}. It 
 includes an evolutionary equation for $\widetilde{\bm{v}}$, which has no
 hydrodynamic counter-part \citep{laibe14}. However, it is
 still possible to improve upon 
 the first-order formulation whilst retaining similarilty to  
 standard hydrodynamics.  
%without having the complexity of the full equations

The full one-fluid equations of motion includes the acceleration term 
\begin{align}
%\left.\frac{D\bm{v}}{Dt}\right|_\mathrm{add} = 
%\bm{a} =  
-\frac{1}{\rho}\nabla\cdot\left(\frac{\rhod\rhog}{\rho}\widetilde{\bm{v}}\otimes\widetilde{\bm{v}}\right) 
\equiv  \frac{1}{\rho}\nabla\cdot \bm{T} 
\end{align}
\citep{youdin05a}, 
where $\bm{T}$ is an effective stress tensor. 
In the terminal velocity
approximation with $\widetilde{\bm{v}} = \tstop 
\nabla P/\rhog$, we have 
\begin{align}
\bm{T} = - \frac{\epsilon\tstop^2}{\rho}\nabla P \otimes \nabla P.
\end{align}
If we define 
\begin{align}
  \bm{g} \equiv \sqrt{4\pi G \epsilon \rho} \frac{\nabla P}{\rho}
  \tstop,\label{effective_g} 
\end{align}
then
\begin{align}
\bm{T} = - \frac{1}{4\pi G} \bm{g}\otimes\bm{g}, \label{effective_gstress}
\end{align}
%{\bf tensor virial theorems?}
which is just the anisotropic part of the  
gravitational stress tensor in self-gravitating hydrodynamics 
\citep{lynden-bell72}.

 However, unlike self-gravity, where $\bm{g}$ is
determined through the Poisson equation; here $\bm{g}$ is directly
related to $P$ (Eq. \ref{effective_g}), which itself obeys an 
evolutionary equation (Eq. \ref{eff_energy}). 
Nevertheless, Eq. \ref{effective_g}---\ref{effective_gstress}
show that correlations in  pressure gradients lead to 
momentum transport of the mixture as a whole. This is similar to Reynolds stresses in
a turbulent disk \citep{balbus99}. In particular, 
outwards angular momentum transport requires $\p_rP\p_\phi P > 0$.

% In a following paper we will include $\bm{T}$ in the one-fluid
% equations and futher explore the analogy between dusty-gas and viscous
% hydrodynamics.

% \subsection{New dust-drag
%   instabilities}\label{dust_as_thermo}

% %{\bf vsi enabled by drag? slipping between dust and gas reduces
%  % effective buoyancy}

% As described in \S\ref{dust_work}, 
% the physical cause of dust-drag instabilities is a phase lag between
% the pressure and density response in a dusty gas\footnote{
% In fact, we expect a phase difference between pressure and the total
% density  whenever $\tstop\neq0$; not just $\tstop\OmK\ll 1$ as 
% considered in this work. This is simply because of the finite
% relaxation time required for dust to respond to gas, and vice versa. 
% An effective energy equation in the form of Eq. \ref{eff_energy} can
% still be obtained for general $\tstop$. The
% heat flux $\bm{F}$ in the cooling function $\mathcal{C}$ is determined
% through solution to the relative velocity $\Delta \bm{v} =
% \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ \citep{youdin05a,laibe14}.  We
% thus expect the `pressure-density lag' interpretation of dust-drag 
% instabilities to be generally applicable.}. 
% This may be used to re-interpret the streaming instability or to
% understand other dusty instabilities. For example,
% \cite{loren15,loren16} have recently discovered a new axisymmetric 
% instability in their two-fluid numerical simulations of dusty
% protoplanetary disks. They find it leads to the formation of
% `toroidal vortices'.

% %give numerical example 
% We have applied our one-fluid stability analysis to find unstable modes  
% similar to that described by \cite{loren15}. We use the same
% parameters reported by them as follows. 
% %we set $p=q=0$, which eliminates VSI and minimizes streaming
% %instabilities as there is no radial pressure gradient at the disk
% %midplane.   
% The total metalicity is $Z=0.01$ and the dust scale-height $H_\mathrm{d} =
% 0.4H_\mathrm{g}$, giving a maximum dust-to-gas ratio
% $\mathrm{max}(\epsilon) = 0.025$. The midplane stopping time is 
% $t_\mathrm{s0}=0.02\OmK^{-1}$. The gas disk parameters are $(p,q,
% \hgas) =(-1.5, -1,0.05)$.  We set $k_x = \pi/H_\mathrm{d}$.

% Fig. \ref{compare_eigenvals_ts0d02} shows the unstable modes found for
% this setup. We find the usual VSI modes, here with $\omega> 3\hgas\OmK,\, s<
% 0.2\hgas\OmK$. However, we also find modes with $\omega<s$. One of these,
% marked by the blue asterisk, has 
% frequency 
% \begin{align*}
% \sigma = (0.3631\ii - 0.1251)\hgas\OmK. 
% \end{align*}
% This growth rate corresponds to $8.8$ orbits, comparable to the $12$
% orbits reported by \citeauthor{loren15}. We visualize this mode in
% Fig. \ref{result2d_loren}. We find this mode 
% persists when $q=0$, so it is not the VSI. 
% Using Eq. \ref{streaming_dispersion} with the above parameters,  
% we also estimate a much smaller growth 
% rate for the streaming instability of 
% $O(10^{-3}\hgas\OmK)$, We thus also rule out
% SI, as did \citeauthor{loren15}. However, the mode disappears with $\tstop=0$, so it \emph{is}
% associated with finite dust-gas drag.

% \begin{figure}
%   \includegraphics[width=\linewidth]{figures/compare_eigenvals_ts0d02}
%   \caption{Unstable modes in a locally isothermal, dusty disk
%     with stopping time  
%     $\tstop=0.02\OmK^{-1}$. Modes with $\omega>3$ are due to
%     VSI. The mode marked by the blue asterisk, visualized in
%     Fig. \protect\ref{result2d_loren}
%     , is associated with finite dust-gas
%     drag. 
%   }\label{compare_eigenvals_ts0d02}
% \end{figure}

% \begin{figure}
%   \includegraphics[width=\linewidth]{figures/result2d_loren.ps}
%   \caption{Visualization of the mode marked by the blue asterisk in
%     Fig. \protect\ref{compare_eigenvals_ts0d02}. 
%     This is a possible eigenmode of the dusty instability discovered by  
%     \cite{loren15} in their numerical simulations. The instability growth timescale and 
%     radial spacing of the `toroidal vortices' are similar to that 
%     reported by \citeauthor{loren15}. This instability is driven by
%     dust-gas drag. The color scale shows the 
%     perturbed  
%     dust-to-gas ratio, $\delta\epsilon$; and the arrows show
%     $\sqrt{\rho}\left(\dd v_x, \dd v_z\right)$. 
%   }\label{result2d_loren}
% \end{figure}

% One possible interpretation of this mode is `drag-driven VSI', to be 
% compared with `cooling-driven VSI' studied by \cite{lin15}. In the
% latter case, buoyancy forces are eliminated by fast
% cooling, allowing vertical shear to destabilize the disk. Now, in an
% isothermal dusty disk if $\tstop=0$ then mixture is 
% effectively adiabatic: dust-induced buoyancy forces are
% strongest. However, having $\tstop\neq0$ allow dust particles to slip
% past the gas, reducing buoyancy. That is, dust-gas drag
% provies an effective cooling.

% We stress that this example only serves to highlight
% the possibility of new axisymmetric dust-drag instabilities in
% stratified disks. A major caveat of the above result is that 
% %the base state is not an exact equilibrium, because dust-settling would occur
% %in practice. (
%  the background dust settling timescale is comparable to the instability
% growth timescale. It is essential to construct
% exact equilibria for a meaningful stability analysis. We defer this
% to future work.

% %also unclear what are "proper" vertical bc for dusty disks

\section{Summary}\label{summary}
In this paper we develop the analogy between isothermal 
dusty-gas and adiabatic, pure gas. The correspondence arises
because drag forces reduce the relative velocity between gas and
dust. In the limit of perfect dust-gas coupling, $\tstop \to 0$,  
 dust is entrained in 
the gas. Then the dust-to-gas ratio $\rhod/\rhog$ is conserved
following the flow. This property is analogous to entropy being conserved
following a adiabatic, pure gas.

For finite drag, $\tstop\neq 0$, the dust content of a 
parcel of the dusty-gas mixture is no longer conserved. The parcel 
can exchange dust particles with neighboring parcels. % owing 
%to dust-gas relative drift. 
This is analogous to heat exchange between a parcel of pure gas and
its surroundings.

We explicitly show that for a fixed gas equation of state, the  
evolutionary equation for the dust-to-gas ratio $\rhod/\rhog$ may be replaced by an 
effective energy equation. This leads to a 
natural definition of the effective entropy of isothermal dusty gas as  
\begin{align*}
  \seff  = \ln{\left(\frac{c_s^2\rhog}{\rhog+\rhod}\right)}.  
\end{align*}
%This allows us to define the buoyancy frequency of an isothermal 
%dusty gas. 
The effect of finite dust-gas friction appears as an energy
source term.  This analogy with standard 
hydrodynamics with cooling/heating allow us to find dusty analogs of gaseous
instabilities, and provide thermodynamical interpretations of  
dust-drag instabilities.

We obtain the equivalent Solberg-Hoiland criteria for the 
axisymmetric stability of strictly isothermal, perfectly-coupled dusty gas.  
Applying this to typical protoplanetary disks, we find that 
the vertical shear associated with dust 
 layers \emph{cannot} lead to axisymmetric  
  instabilities, however thin the dust layer is.    
Instead, sharp radial  edges in the dust-to-gas ratio could destabilized the
disk, as these imply sharp gradients in the disk's effective entropy
profile. Alternatively, if the dust is vertically well-mixed, then any
radial gradient in $\rhod/\rhog$ can destabilize the disk.

 % {\bf presumably this limits how sharp dust rings can be}

We  apply our thermodynamic framework to interpret
the streaming instability \citep[SI, ][]{youdin05a,jacquet11} and to generalize the vertical shear
instability \citep[VSI, ][]{nelson13,lin15} to dusty disks. We
explicitly show in SI the evolution of gas
pressure lags behind dust density. This is in fact a general property of 
instabilities driven by dust-gas drag. 
It takes a finite time for the gas to respond to
the dust motion. A lag implies there exists a time interval where 
the gas pressure of a parcel of the dusty-gas mixture is increasing whilst dust is already being expelled.  
The dusty gas then does positive work that amplifies 
oscillations. This interpretation is analogous to stellar pulsational
instabilities \citep{cox67}.

For the VSI we find dust-loading is generally stabilizing. In our disk models 
dust-loading does not affect VSI growth rates significantly, but
meridional motions may be suppressed where the dust-induced
vertical buoyancy dominates over vertical shear, consistent with our
previous study \citep{lin15}. Since the dust-induced buoyancy forces
increase away from the midplane, we find dust-loading can stabilize
`surface modes'  of the VSI, that would otherwise have the largest
growth rates.

We also discuss future applications of our thermodynamic framework to 
study dusty protoplanetary disks. Because isothermal dusty gas has an
effective entropy, we suggest that purely hydrodynamic processes, such
as the Rossby Wave Instability \citep{li00} or the `Zombie Vortex
Instability' \citep{marcus15}, where entropy plays a role, could have dusty counter-parts.   
Furthermore, hydrodynamic instabilities driven by thermal cooling, such as VSI in
stably-stratified disks \citep{lin15} or the `Convective
Overstability' \citep{klahr14, lyra14},  may also find dusty analogs 
because finite dust-gas drag is equivalent to a heat sink/source.

%In particular, we emphasize the 
%physical origin of any instability driven by strong dust-gas
%drag is a phase lag between the pressure and total density evolution
%of the dusty gas mixture. This leads to growing modes because 
%positive work is done during an oscillation cycle, provided by
%dust-gas drag, which is equivalent to a `heat' 
%source.

%We demonstrate a potentially new dust-drag instability in
%stratified disks that may explain

%We also applied
%our framework to find unstable modes in dusty, stratified disks
%similar to that reported recently by \cite{loren15}.