### 2017-09-12 response.txt
first draft of revision 1 and response to referee

We thank the referee for a prompt report, which helped to improve the
clarity of the manuscript. We have made revisions in bold, and further
explain them below.

Major points:

* The section 2 is interesting as it relates the work done by
  pressure, when not in phase with density, to the growth of
  oscillations. Although I understand that it is a link between the
  pure gas context and the dusty flow case (because the drag force
  acting on the gas can dephase density), I think the authors should
  approach the point later in the paper. In my opinion, it would be
  more adequate to approach it at the end of or after the actual
  section 3 (single fluid description of dusty gas). Then it is
  relevant to make the parallel with the pure gas dynamics. It is also
  a way to explain the analogy that appears in parag. 3.3 "The
  physical reason for this analogy..."

  We appreciate the referee’s suggestion that section 2 might make
  sense if moved.  With respect, our choice for the placement of this
  section is meant for readers who are lack the technical expertise of
  the reviewer. Specifically, our goal is to summarize a main result
  and a key physical insight, which motivates the technical
  derivation of the dusty/adiabatic gas analogy. We added a new
  paragraph at the beginning of section 2 to state its purpose. 
  We hope section 2 motivates further reading and of course any reader
  can revisit  section 2 (section 4.4 refers back to it).

While the referee is certainly correct that the single fluid analogy
needs to made, we have attempted to explain the concept in the
introduction. In any event, we hope the referee will agree to some
authorial prerogative in the order of presentation.

* In section 6 on the numerical results, in particular parag. 6.5 and
  6.6, I wonder which growth rates are obtained with different radial
  mode numbers. It would be an advantage to see other values as in Fig
  8. In 6.5 kx is 30, does the instability change with smaller, larger
  modes? It is worse in 6.6, where kx is very large, 1800. Is it
  relevant to have such a high radial mode, that could barely be
  resolved in non linear numerical simulations? Finally, in a general,
  I think that a more physical summary of the results of the section 6
  should be provided. It is almost limited now to a description of the
  results.

  For section 6.5 we have updated Fig. 14 to include cases with
  different kx. All growth rates remain the same order of
  magnitude. The effect of a radial varation in rhod/rhog is also
  similar between different kx. However, surface modes only appear at
  high kx (as found for pure gas VSI, Barker & Latter, 2017)

  For section 6.6 we also updated Fig. 15 with more kx values. 
  We find to obtain instability with smaller kx, one needs a more
  rapidly-varying dust-to-gas ratio. We do not dwell 
  on this case because we have already shown, using the more general
  analyses in section 4.3 and 5.2

  Note that while the original case with high kx=1800 may not be  
  resolved in global simulations, it can be resolved in local shearing
  boxes. Simulations of SI in fact consider even larger
  kx = O(10^3/h), where h= H/R is the gas aspect-ratio. See, Bai &
  Stone (2010) and Yang & Johansen (2017), for example.

  In section 8 we extend the summary on the results of section
  6. We discuss possible implications of our results in 
  realistic disks with finite stopping times. In particular, we
  suggest the criterion for particles to settle against the VSI.

* The section 7 'Discussion' is for the moment a list of common
  instabilities that could be triggered by the effect of the drag in
  the framework of the one-fluid approximation. Even if it is very
  interesting, it may be too speculative; I would not limit the
  discussion only to this. The authors may first discuss more about
  the limitations/assets of their approach based on the numerical
  results obtained in section 6, which will give arguments to the
  possible numerical implementation of the method (see parag. 7.3). In
  particular, what is the impact of the terminal velocity
  approximation? How much reliable the results and the accuracy of the
  approach are with a finite St number? For example, what do you
  expect for a typical value of 0.01-0.1 (dust or pebble), for which
  it is known that the terminal velocity is inappropriate?

  We have added a new section 7.1 on the application and limitations
  of the one-fluid framework.

  The terminal velocity approximation from the requires ts to be small
  compared to other timescales in the problem. Since 1/Omega is the
  typical `fast' timescale for disks, we generally require ts*Omega <<
  1, but the exact critical value is problem-dependent.

  Fig. 4 showsthat the one-fluid framework can gives accurate
  streaming instability growth rates in dust-rich disks even for
  ts*Omega = 0.1. However, the dust-poor `linB' mode in Table 1 (also
  with ts*Omega=0.1) is not accurately captured by the one-fluid framework. 
  This suggest that the terminal velocity approximation is not the
  only condition required for the one-fluid equations to give accurate
  results. In the case of the streaming instability, it seems that one
  also requires rhod > rhog.

* In the Appendix: parag. D.2.2, the authors check if two assumptions
  done during the calculations are valid (eq D11 and eq D16). But it
  seems that they compute the values with the results obtained after
  using these approximations. Please clarify

  We retitled this section to `Consistency check' since, as the
  referee pointed out, we simply checked that the assumptions made to
  obtain the explicit solutions are self-consistent. The
  approximations are ultimately justified by comparison to numerical
  solutions in Fig. 4.

* Last point: the authors should re read carefully the article, indeed
  some sentences are not completly made agree with their later
  modifications.

  We have proof-read the new manuscript.

Minor points:

* parag 4.4.1 Can you explain briefly in the manuscript how to obtain
  the eq 33. ?

  Additional description (Eq. 33) is now included.

* parag. 5.1.2 Can you clarify the text related to eq 47, one cannot
  see clearly which terms are the second and the third. Maybe one can
  say : "The first term in brackets is usually stabilizing ... "

  This is clarified.